\pdfbookmark{Acknowledgements}{Acknowledgements}

\addchap*{Acknowledgements} %instead of chapter, so it won't be numbered

\begin{doublespacing}
I would like to express my deepest appreciation to Dr. Oren Shriki. Thank you for allowing me to travel my own path in the Garden of Science you've helped cultivate. Your heartfelt mentorship and support during my two years at the Computational Psychiatry Lab will never be forgotten.

This endeavor would not have been possible without Dr. David Tolpin. Thank you for advising, teaching and collaborating with me throughout my journey. I learned from you how to articulate basic scientific questions, and how to effectively apply myself to answering them.

Special thanks to my teachers and administrative staff at Ben-Gurion University of the Negev, and especially the Computer Science, Mathematics and Cognitive Science departments. You have provided me all of the educational and physical resources I could ask for to realize my academic dreams.

I could not have undertaken this journey without my lab mates. Thank you for giving me inspiration, fruitful arguments and being there for me in times of need: Aviv Dotan, Miriam Guendelman, Daniel Polyakov, Lahav Foox, Miriam Dissen, Asaf Harel, Shay Ben-Sasson, Adi Nissim, Yarden Ben-Horin, Sagi Furman, Ofer Avin, Ophir Almagor, Anat Shkedy-Rabani, Or Rabani.

% % \vspace{2em}
% % \thesisAuthor{}
To Miriam Guendelman, again, for your insightful remarks and editing suggestions. I hope to be able to pay you back soon.

To the Mayo Foundation for Medical Education and Research, for granting me permission to use your art (figure \ref{fig:c3bsle:caninedb}).

To my family - grandparents, parents, siblings, fiancé and others. Your inspiring achievements, loving support and affirmative relationships, at all times, are the main reasons why I'm here today. If it were not for you, this thesis would not exist.

\end{doublespacing}
